Stack Safety is a Security Property

### 2021-04-10 tex/main.tex
Redid section 3

\begin{figure}[b]
  \centering
  \begin{subfigure}{.45\textwidth}
\begin{verbatim}
int stash;

int main() {
0:             int x = 42, y = 0;
1[P(x)],2[C]:  f(&x);
3:             x = 0;
4[P(y)],5[C]:  f(&y);
6[R]:          return 0;
}

void f(int* a) {
10:  if(!stash)
11:    stash = a;
12:  *stash = 5;
13:  return;
}
\end{verbatim}
\end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \begin{tabular}{l l}
      {\tt 2} &
      \memoryaddrs{12em}
      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
      \memory{2}{\unsealc}[{\makebox[0pt]{Unsealed}}]
      ~$\cdots$ \\
      {\tt 10} &
      \memoryaddrs{17em}
      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{2}{\unsealc}
      ~$\cdots$ \\
      {\tt 3} &
      \memoryaddrs{12em}
      \memory{3}{\unsealc}
      ~$\cdots$ \\
      {\tt 5} &
      \memoryaddrs{12em}
      \memory{1}{\unsealc}
      \memory{1}{\mainpassc}
      \memory{1}{\unsealc}
      ~$\cdots$ \\
      {\tt 10} &
      \memoryaddrs{17em}
      \memory{1}{\mainsealc}
      \memory{1}{\mainpassc}
      \memory{2}{\unsealc}
      ~$\cdots$ \\
\end{tabular}
\end{subfigure}

@@

Up til now we have assumed that function arguments and returns are passed solely in registers.
This is an oversimplication, which we will now correct. There are important stack safety concerns
when passing arguments on the stack, as seen in example \ref{fig:passing}. In this example,
{\tt main} passes a reference to {\tt x}, but then clears {\tt x} and instead passes one to
{\tt y}. Before each call, when the address of a variable is taken, we add a new annotation
\(P(x)\) or \(P(y)\) identifying the memory to be passed, and our stack model marks it in
magenta. \apt{Why are there two line numbers assoicated with the two calls?}
Then at the call, the parts of the caller's frame that are not passed are sealed, as usual. Passed
memory also tracks which depth owns it, and it may only be accessed by the immediate callee
(at its depth plus one).

\apt{This needs to be talked through in more detail.}
But because {\tt f} stashes the original pointer to {\tt x}, it will overwrite {\tt x} while
{\tt x} is sealed, an integrity violation. This fits with the intuition of passing by reference,
that the passed variable might be read or written during the call, not that it is permanently
open to be read or written by all subsequent calls. This also applies to passing large objects
by value in the stack: the caller allocates them, and the relevant instructions are annotated to
indicate passing.

We do not model cases where permanent sharing is desireable, because they are uninteresting
and hard to motivate: we would expect to treat such sharing as completely unprotected, and
perhaps attach a separate memory safety policy to add some protection to such variables.

>>>>

\begin{figure}
  \centering
  \begin{subfigure}{.45\textwidth}
\begin{verbatim}
int stash = 0;

int main() {
0:         int x, y;
1[P(x)C]:  f(x);
2:         x = 0;
3[P(y)C]:  f(y);
4[R]:      return x;
}

void f(int &a) {
10:    if(!stash)
11:      stash = &a;
12:    *stash = 5;
13[R]:   return;
}
\end{verbatim}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \begin{tabular}{l l}
      {\tt 10} &
      \memoryaddrs{17em}
      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$ \\
      {\tt 11} &
      \memoryaddrs{17em}
      \memory{1}{\mainpassc}
      \memory{1}{\mainsealc}
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{5}
      \\
      {\tt 3} &
      \memoryaddrs{12em}
      \memory{4}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 10} &
      \memoryaddrs{17em}
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 11} &
      \memoryaddrs{17em}
      \memory{1}{\mainsealc}
      \memory{1}{\mainpassc}
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{\bf 5}
\end{tabular}
\end{subfigure}

@@

Up til now we have ignored function arguments, since our simplest model places them in
registers. This is an oversimplication, which we will now correct, starting with the more
interesting case: call-by-reference. There are important stack safety concerns
when passing arguments by reference, as seen in example \ref{fig:passing}. In this example,
{\tt main} calls {\tt f} with a reference to {\tt x}, but then clears {\tt x} and instead
passes a reference to {\tt y}. We would expect that the first call to {\tt f} might modify
{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
{\tt x} in the second call.

At each call, we add a new annotation \(P(x)\) or \(P(y)\) identifying the memory to be
passed, and our stack model marks it in magenta. Then the caller's frame that are not
passed are sealed, as usual. Just as we marked sealed memory with the depth of its owner,
we do the same to passed memory. We keep track of the current depth of the active function
next to the stack pointer in the diagram. A passed address may only be accessed when the
current depth is one higher than the one it is sealed with.

So at the initial call, {\tt x} is passed, and {\tt f} is free to write to it. Then upon
return, {\tt main}'s frame is unsealed. At the second call it is {\tt y} that is passed,
and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
a future call with no arguments passed by reference would also violate integrity if it wrote
to either variable.

Passed addresses are also excluded from confidentiality: at the start of each call, we vary
only stack addresses that are not passed. This is important, as we use the passing label
for those arguments that are passed by value on the stack, as well. We might instead
treat them as unsealed, but then we would treat them as confidential like the rest of the stack.

We do not model permanent sharing, in which a variable's address is taken and it might be
passed far afield before use, because this situation is uninteresting and hard to motivate
as an instance of stack safety. We would expect to treat such sharing as completely
unprotected, and perhaps attach a separate memory safety policy to add some protection
to such variables.
